---
title: "Introduction to options.data"
author: "Jason Taylor"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to options.data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
options(tibble.print_min = 4L, tibble.print_max = 4L)
library(dplyr)
library(data.table)
```

When creating tidy options data you must:

* Gather the raw options data (this package requires iVolatilty source data)

* Remove dates where the market was closed

* Add variables to the dataset used in studies ie. (rsi, ivrank, dte)

The options.data package makes these steps fast and easy:

* It provides simple functions that correspond the most common issues with raw option data.

* Requires the TTR package so that you can add additional indicators easily.

This document introduces you to options.data's functions and shows you how to apply them.

## Raw options data: 

Sample of the data retrieved from iVolatility.com read in from the raw .csv files

```{r, echo=FALSE, results='asis', warning=FALSE, message=FALSE, error=FALSE}
library(options.data)
knitr::kable(raw_example[, 1:10])
knitr::kable(raw_example[, 11:22])
```

As you can see the dataset is not quite ready for use in studies:

* We can safely remove the exchange, adjusted stock price, style, and asterix columns

* We also need to add the rsi_14, iv_rank_90, and iv_rank_252 values to the dataset

## Functions included in the package:

* `tidy_options()`

* `first_day_month()`

* `first_day_week()`

* `daily()`

* `market_closed()`

* `third_friday()`

* `earnings_dates()`

## Create tidy options data with `tidy_options()`

`tidy_options()` creates the final tidy output data.frame.  The first arguement is
the name of the underlying you are processing, for example "XLE".  The second and
third arguements are the folder holding the raw options .csv files, and the iv file
with the root being your working directory. 

Example call:
```{r, eval=FALSE}
tidy_options("XLE", "data/raw_files/options", "data/volatility/vx.xle.daily.prices.RData")
```

Output is saved to "data/options/`symbol`.options.RData"

Sample output: 
```{r, echo=FALSE, results='asis', warning=FALSE, message=FALSE, error=FALSE}
knitr::kable(processed_example[, 1:7])
knitr::kable(processed_example[, 8:16])
knitr::kable(processed_example[, 17:23])
```

## Create data.frames of dates to open trades:

_each function takes the arguements `start_yr` and `end_yr`_

* `first_day_month()` creates data.frame of first day in month adjusted for
market closed dates. `monthly.RData` is included in the package with start and 
end dates of 2010 - 2020.

* `first_day_week()` creates data.frame of first day in week adjusted for
market closed dates. `mondays.RData` is included in the package with start and 
end dates of 2010 - 2020.

* `daily()` creates data.frame with all days the market was open. 
`daily.RData` is included in the package with start and 
end dates of 2010 - 2020.

## Market closed:

`market_closed` creates data.frame of all dates the market was closed from 
`start_yr` to `end_yr` `market_closed.RData` is included in the package with start and 
end dates of 2010 - 2016.

## Calculate monthly expiration dates `third_friday()`

`third_friday` calculates the monthly expiration dates for the arguements 
`start_yr`, `end_yr` which is then used in studies to limit opening trades to 
monthly expirations. These values are not exported as a dataset but rather added 
to the processed option dataset.

## Earnings dates:

`earnings_dates()` creates a data.frame of the earnings dates from csv files that
need to be updated manually as the time window of data expands.
`earnings.RData` is included in the package with start and 
end dates of 2010 - 2016.
